Solar wind predictions for the Parker Solar Probe orbit

### 2017-08-10 paperMVVB/sections/content.tex
implementing VBedits

\color{Green4}
%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1962 \citep{Neugebauer1966}.

%in situ spacecraft
The idea of a space mission flying through the solar corona \hl{dates back to the founding year of NASA in 1958 (ref.).} Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in the case of Voyager~1 as far away as \SI{138}{\au}\footnote{\url{https://voyager.jpl.nasa.gov/}}, having even left the heliospause into interstellar space \citep{Gurnett2013}. Ulysses was the unique probe that orbited the Sun out of the ecliptic plane and thus could retrieve solar wind measurements from the poles of the Sun \hl{(ref.)}. So far, Helios~2 made the nearest in situ solar wind measurements ever at a distance as close to the Sun as \SI{0.29}{\au}, closely followed by Helios~1 with \SI{0.31}{\au} \citep{Rosenbauer1977}.

%Solar Probe Plus mission
The Parker~Solar~Probe, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia with 9.86~solar radii, i.e. \SI{0.0459}{\au}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24 orbits with perihela inside \SI{0.25}{\au} \citep{Fox2015}.

\color{Red2}
%PSP objectives and instruments
The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” (ref. NASA STDT report, \citet{Fox2015}). To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields, SWEAP for the measurements of flux of electrons, protons and alphas, IS\sun{}IS for the measurement of solar energetic particles and WISPR for the measurement of coronal structures (ref. to each instrument paper).\\

%WISPR
The Wide field Imager for Solar Probe (WISPR) will contribute to the science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures, e.g., streamers and pseudostreamers, equatorial coronal holes, for determining the roles of turbulence, waves, and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps, of SEP-producing shocks and their CME drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.\\

%CGAUSS
CGAUSS (Coronagraphic German And US Solar Probe Survey) is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR). One of the objectives of the CGAUSS project is to extrapolate the solar wind environment down to the closest perihelion of \SI{9.86}{R_s} distance to the Sun in order to optimize the WISPR and PSP science operations. To achieve that goal this study uses in situ solar wind data from the US-German Helios~1 and 2 space probes and \SI{1}{\au} data from various satellites compiled in the NASA/GSFC OMNI solar wind database.\\

\color{Green4}
This paper examines the variations of the bulk solar wind frequency distributions, considers solar activity and extrapolates these distributions to the near-Sun region.

\color{black}
%definition of sw environment
\section{Solar wind environment}

\color{Green4}
%solar wind parameters
As a baseline we treat the solar wind primarily as a proton plasma, because the average helium abundance is only about \SI{4.5}{\percent} and in slow wind at solar cycle minimum even less than \SI{2}{\percent} \citep{Feldman1978,Schwenn1983,Kasper2012}. Neglecting heavier ions, the electron and mass density can then \hl{be derived accordingly.}

The characteristic behavior of a magnetized plasma is determined by its \textit{density}, \textit{temperature} and \textit{magnetic field strength}. Furthermore, the bulk flow \textit{velocity} is the parameter which makes the plasma a 'wind'. For this study we define the solar wind environment through these four major solar wind parameters. Quantities like flux densities, mass flux and plasma beta can directly be derived from those four parameters.

Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966}\hl{, Schwenn (19xx), McComas et al. (19xx)).} Fast streams are found to originate from coronal holes as Ulysses polar measurements revealed \hl{(ref.). The origin of slow wind is a subject of controversial discussions because several sources are closed magnetic structures in the solar corona, such as Schwenn (19xx) has even reported the possibility of four types of solar wind depending on the phase of the solar cycle/activity.}

The solar wind velocity is the defining parameter of the two types of solar wind. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[see][p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}.

Their occurrence frequency varies strongly with solar activity and their interactions lead to phenomena such as \color{red}
stream interaction regions (SIRs) and for quasi-stationary coronal source regions to co-rotating interaction regions (CIRs, ref. Buch über CIRs).

\color{Green4}
Embedded in the slow and fast solar wind streams are coronal mass ejections (CMEs) \hl{(ref.)}. Their frequency in near \SI{1}{\au} measurements varies between almost zero during solar cycle minima \hl{up to a daily rate of about 0.5 during times of solar maximum} \citep{Richardson2012}.

%frequency distribution data
Since one cannot know which specific solar wind type or structure PSP will encounter at a given point in time during its mission, we extrapolate the parameters’ probability distributions from existing solar wind measurements.

\color{black}
Our approach is to get an analytical representation of the frequency distributions' shapes (Sect.~\ref{sec:frequency_distribution}), their solar activity dependence (Sect.~\ref{sec:solar_activity_variations}) and their solar distance scaling (Sect.~\ref{sec:solar_distance_dependency}). We get the parameters' frequency distributions and solar activity dependence from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles and their distance dependency from solar wind measurements of more than half a solar cycle, covering more than two third of the distance to the Sun (\SIrange{0.29}{0.98}{\au}).%0.68~au\\

From the combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions we build a general model representing the solar activity and distance behavior of all four solar wind parameter frequencies (Sect.~\ref{sec:general_solar_wind_model}).

This general model is then fed with a SSN prediction and extrapolated to PSP's planned orbital positions (Sect.~\ref{sec:model_extrapolation_to_psp_orbital_time_and_position}).\\

%lognormal fitting
\section{Frequency distributions}
\label{sec:frequency_distribution}

This section looks at the solar wind parameters' frequency distributions which we extract from the in situ OMNI data set. We determine adequate fit function types and evaluate how suitable they are to represent the frequencies' shapes.

\subsection{OMNI frequency data}
\label{sec:omni_frequency_data}
The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on solar activity. Hence, for deriving general frequency distributions of the solar wind parameters, averaging over long-term solar wind variations is needed. This requires a distance-independent data set covering multiple solar cycles. The abundance of near-Earth hourly OMNI data is well suited for this task, because it spans almost five solar cycles.

%about OMNI data
This OMNI~2 data set \citep{King2005} combines solar wind plasma and magnetic field data and for this study not important energetic proton fluxes, geomagnetic and solar indices. Because it covers decades, the near-Earth solar wind data is composed of intercalibrated multi-spacecraft data which is time-shifted to the nose of the Earth's bow shock.

%data source
The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
The hourly data of the whole time range up to \mbox{2016-12-31} is the basis for the frequency fits. The data starts in \mbox{1963-11-27}, but the temperature data not before \mbox{1965-07-26}. The data coverage of the parameters is between \SIrange{67}{74}{\percent}, which adds up to about 36--40~years in total. This plethora of data is well suited for our task.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\

%hourly data
We note that the use of higher time resolution instead, would not significantly change our results. The frequency distributions of the also available minutely OMNI data set are almost congruent with the hourly---they only differ slightly at their extreme ends.

%data binning
We specify bin sizes considering the individual maximal parameter ranges and the OMNI data precision. Especially for the density and temperature we choose their bin sizes such small that their distributions' peaks can be resolved (the peaks are at their lower end). We set the individual bin sizes to \SI{0.5}{nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature.

Next, we look for a suitable fit function for the resulting histogram shape of the solar wind parameters' frequency distributions.

\subsection{Lognormal fitting}
%choice of distribution function---why lognormal function?
Obviously all possible values for the four parameters are positive. This hints to the supposition that they are lognormally distributed, as many positive natural quantities conform to lognormal distributions. Its probability density function is described by a lognormal function.
%fitting method
Therefore we use a lognormal function as the fit function in the process of the least squares regression fitting.
%fit function
The lognormal function

@@

depends on the location $\mu$ and the shape parameter $\sigma$. Changes in $\mu$ affect both the horizontal and vertical scaling of the function whereas $\sigma$ influences its shape. The distribution's median $x_\text{med}$ and mean $x_\text{avg}$ (average) positions are easier to interprete and can directly be calculated from $\mu$ and $\sigma$:

@@

The values of $x_\text{med}$ and $x_\text{avg}$ obtained from fitting the solar wind frequency distributions are listed in Table~\ref{tab:lognormal_fit_parameters}.

@@

From visual inspection, the resulting curves match well with the shape of the magnetic field strength, density and temperature distributions (Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}). However, for the velocity the fit function seems insufficient in describing its more complex shape, especially at its peak position and the faster end of the distribution. Its sum of absolute residuals (SAR) between data and fit is almost three times larger than those from the other parameters (Table~\ref{tab:lognormal_fit_parameters}). They can be compared, because the area of probability density functions is unity.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits. The histograms have bins of 0.5~nT, 10~km/s, 1~cm$^{-3}$ and 10\,000~K and are based on the hourly OMNI data set. The fit's median and mean values are indicated as well. The insets only have zoomed-in frequency axes, their x-axes stay the same.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
To reach a better fit result for the velocity we change the fit function. We do not want to abandon the well-founded application of the lognormal function. However, it is reasonable to assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach promises better fit results, which is why we combine two lognormal functions (\ref{eq:single_lognormal_fit_function}), bearing the disadvantage of more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution gives the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$), which are also listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0
\end{align}
respectively.\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{The velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
For this reason we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

\subsection{Power law lognormal fitting}
To retrieve the frequency distributions for every \SI{0.01}{\au} solar distance bin, we choose the same solar wind parameter binning as with the OMNI data (Sect.~\ref{sec:omni_frequency_data}).
% Helios histogram bin sizes for mean of frequency distribution (at specific solar distance):\\
% $B$: bin size 0.5~nT, min 337, mean precision: 0.000545\\
% $v$: bin size 1~cm$^{-3}$, min 497, mean precision: 0.00449\\
% $n$: bin size 10~km/s, min 497, mean precision:  0.00449\\
% $T$: bin size 10\,000~K, min 497, mean precision: 4.49--44.49\\

>>>>

%solar wind beginnings
From observations of cometary tail fluctuations \citet{Biermann1951} inferred the presence of a continuous flow of particles from the Sun. With his theoretical solar wind model \citet{Parker1958} formulated the existence of the solar wind even before the first satellites measured it in situ in 1962 \citep{Neugebauer1966} \hl{+ ref. to Russian measurements}.
%in situ spacecraft
The idea of a space mission flying through the solar corona dates back to the founding year of NASA in 1958 \citep{McComas2008}. Since then several space missions have measured the solar wind in situ at a wide range of heliocentric distances, in the case of Voyager~1 as far away as \SI{138}{\au} in July 2017\footnote{\url{https://voyager.jpl.nasa.gov/}}, having even left the heliospause into interstellar space at a distance of \SI{121}{\au} \citep{Gurnett2013}.
Until today various spacecraft have provided a wealth of solar wind measurements near Earth’s orbit, with Wind \hl{(ref.)}, ACE \hl{(ref.)} and DSCOVR \hl{(ref.)} still orbiting around the L1 point \SI{1.5}{million \km} ahead of Earth in the sunward direction. Additional measurements at other distances were provided by planetary missions to Venus and Mercury, such as PVO \hl{(ref.)} or MESSENGER \hl{(ref.)}. Ulysses was the first probe that orbited the Sun out of the ecliptic plane and thus could measure solar wind at polar latitudes \citep{McComas1998}. The nearest in situ solar wind measurements to date were obtained by the Helios mission. The in 1974 launched Helios~1 spacecraft reached distances of \SI{0.31}{\au}, Helios~2 launched two years later and approached the Sun up to \SI{0.29}{\au} \citep{Rosenbauer1977}.

%Solar Probe Plus mission
The NASA Parker~Solar~Probe\footnote{\url{http://parkersolarprobe.jhuapl.edu/}} (PSP), formerly Solar~Probe~Plus, with a planned launch date in mid 2018, will reach after six years in 2024 its closest perihelia at a distance of 9.86~solar radii (\Rsun), that is, \SI{0.0459}{\au} \citep{Fox2015}. This distance will be achieved through seven Venus gravity assists with orbital periods of 88--168~days. In its prime mission time 2018--2025 PSP provides 24~orbits with perihelia inside \SI{0.25}{\au} \citep{Fox2015}. Even its first perihelion, 93~days after launch in 2018, will take PSP to an unprecedented distance of \SI{0.16}{\au} (\SI{35.7}{\Rs}). {\color{Red2} In comparison, the ESA Solar Orbiter mission with a planned launch in 2019 will have a closest perihelion of \SI{0.289}{\au} \hl{(ref.)}.}

%PSP objectives and instruments
The key PSP science objectives are to “trace the flow of energy that heats and accelerates the solar corona and solar wind, determine the structure and dynamics of the plasma and magnetic fields at the sources of the solar wind, and explore mechanisms that accelerate and transport energetic particles” as stated in \citet{Fox2015}. To achieve these goals, PSP has four scientific instruments on board: FIELDS for the measurements of magnetic fields and AC/DC electric fields \citep{Bale2016}, SWEAP for the measurements of flux of electrons, protons and alphas \citep{Kasper2016}, IS\sun{}IS for the measurement of solar energetic particles \citep{McComas2016} and WISPR for the measurement of coronal and inner heliospheric structures \citep{Vourlidas2016}.

%CGAUSS and WISPR
The study presented in this paper is undertaken in the Coronagraphic German And US Solar Probe Survey (CGAUSS) project, which is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR). WISPR will contribute to the PSP science goals by deriving the 3D structure of the solar corona through which the in situ measurements are made to determine the sources of the solar wind. It will provide density power spectra over a wide range of structures (e.g., streamers, pseudostreamers and equatorial coronal holes) for determining the roles of turbulence, waves and pressure-balanced structures in the solar wind. It will also measure the physical properties, such as speed and density jumps of SEP-producing shocks and their CME drivers as they evolve in the corona and inner heliosphere \citep{Vourlidas2016}.

In order to help optimize the WISPR and PSP preplanning of the science operations knowledge of the expected solar wind environment is needed. For this purpose the solar wind environment is extrapolated down to the closest perihelion of \SI{9.86}{R_s} distance to the Sun using in situ solar wind data from the Helios probes and near \SI{1}{\au} data from various satellites compiled in the NASA/GSFC OMNI solar wind database.\\

%definition of sw environment
%section{Solar wind}

Generally, two types of solar wind are observed in the heliosphere, slow and fast streams \citep{Neugebauer1966}\hl{, Schwenn (19xx)}. Slow solar wind has typical speeds \SI{<400}{\km\per\s} and fast solar wind has speeds \SI{>600}{\km\per\s} \citep[p.~144]{Schwenn1990}. Their different compositions and characteristics indicate different sources and generation processes \citep{McGregor2011a}. Fast streams are found to originate from coronal holes as confirmed by Ulysses out-of-ecliptic measurements \citep{McComas1998}. The source of slow wind and its eventually different types (\hl{Schwenn19xx}), is still a subject of controversial discussions because several scenarios are possible to explain its origin from closed magnetic structures in the solar corona, {\color{red} such as ...intermittent reconnection at the top of streamers and pseudostreamers, coronal hole boundaries, ... (\hl{ref.}).} The occurrence frequency of these slow and fast streams varies strongly with solar activity and their interactions lead to phenomena such as stream interaction regions and for quasi-stationary coronal source regions to co-rotating interaction regions \citep{Balogh1999}.

Embedded in the slow and fast solar wind streams are transient flows of coronal mass ejections (CMEs), the faster ones driving shock waves ahead \hl{(ref.)}. Their rate follows the activity cycle and varies in near \SI{1}{\au} measurements between only one CME every couple of days during solar cycle minima up to multiple CMEs observed over several days at times of solar maxima, that is, the CME-associated flow share of the solar wind raises from about \SI{5}{\percent} up to about \SI{50}{\percent} \citep{Richardson2012}.

It is not known which specific solar wind type or structure PSP will encounter at a given time during its mission, therefore we extrapolate the probability distributions of the major solar wind parameters from existing solar wind measurements and take solar cycle dependencies into account.

%solar wind parameters
As a baseline we describe the solar wind environment through the key quantities of a magnetized plasma: \textit{density}, \textit{temperature} and \textit{magnetic field strength}. Furthermore, the bulk flow \textit{velocity} is the defining parameter of the two types of solar wind. Solar wind quantities, like flux densities, mass flux and plasma beta, can directly be derived from these four parameters.

Our approach is to obtain analytical representations of the shapes of the solar wind parameter’s frequency distributions in Sect.~\ref{sec:frequency_distribution}, of their solar activity dependence in Sect.~\ref{sec:solar_activity_variations} and of their solar distance scaling in Sect.~\ref{sec:solar_distance_dependency}. The solar wind parameters’ frequency distributions and solar activity dependence is derived from near-Earth solar wind and sunspot number (SSN) time series with a duration of almost five solar cycles. Their distance dependency is derived from Helios solar wind measurements covering more than two third of the distance to the Sun and more than half a solar cycle. From combination of the obtained frequency distributions, SSN dependence functions and solar distance dependence functions a general solar wind model is build in Sect.~\ref{sec:general_solar_wind_model}, representing the solar activity and distance behavior. Finally, this empirical model is fed with a SSN prediction and extrapolated to PSP's planned orbital positions in Sect.~\ref{sec:model_extrapolation_to_psp_orbital_time_and_position}.

%lognormal fitting
\section{Frequency distributions of the solar wind parameters}
\label{sec:frequency_distribution}

The solar wind parameters are highly variable, due to short-term variations from structures like slow and fast wind streams, interaction regions and CMEs, whose rate and properties depend on the phase of the solar activity cycle. Hence, for deriving characteristic frequency distributions for the solar wind parameters, measurements over long-term time spans are needed. The abundance of the near-Earth hourly OMNI data set is ideally suited for this purpose, because it spans to date almost five solar cycles.

%about OMNI data
The OMNI~2 data set \citep{King2005} combines solar wind plasma and magnetic field data collected by various satellites since 1963. This intercalibrated multi-spacecraft data is time-shifted to the nose of the Earth’s bow shock. The data is obtained from the OMNIWeb interface\footnote{\url{http://omniweb.gsfc.nasa.gov/}} at NASA's Space Physics Data Facility (SPDF), Goddard Space Flight Center (GSFC).
%OMNIWeb Data Documentation: http://omniweb.gsfc.nasa.gov/html/ow_data.html
% Acknowledgement to the SPDF OMNIWeb database as the source of data used in publications is requested: "The OMNI data were obtained from the GSFC/SPDF OMNIWeb interface at http://omniweb.gsfc.nasa.gov". Further, for recent years when few sources (IMP 8, Wind, ACE, Geotail) contributed to OMNI, it would be appropriate to also cite the PI's who provided the data to OMNI. Copies of preprints or reprints of OMNI-based publications sent to Natalia Papitashvili (address below) would be appreciated for tracking purposes.
% The best citable reference to OMNI data is J.H. King and N.E. Papitashvili, Solar wind spatial scales in and comparisons of hourly Wind and ACE plasma and magnetic field data, J. Geophys. Res., Vol. 110, No. A2, A02209, 10.1029/2004JA010804.
%
%data time range
In this study the whole hourly data until 31 December 2016 is used, starting from 27 November 1963 (for the temperature from 26 July 1965). The data coverage of the parameters is in the range \SIrange{67}{74}{\percent},  corresponding to a total duration of 36--40~years.
%OMNI 1963-01-01 data
%B 1963-11-27 14:30
%V 1963-11-27 12:30
%N 1963-11-27 12:30
%T 1965-07-26 00:30
%1963-01-01~00:00--2016-12-31~23:30 all
%19724 d; 54 y\\
% B: 73.91 \%		39.91 y\\
% V: 73.07 \%		39.46 y\\
% N: 70.29 \%		37.96 y\\
% T: 66.73 \%		36.03 y\\
%
%hourly data
It should be noted that a test-comparison of hourly averaged with higher time resolution data for the available shorter time span 1981--2016, did not show significant differences in our results.

%data binning
The frequency distributions of the solar wind magnetic field strength, proton velocity, density and temperature are shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. According to the OMNI data precision and maximal parameter ranges we specify the bin sizes to \SI{0.5}{\nT} for the magnetic field strength, \SI{10}{\km\per\s} for the velocity, \SI{1}{\per\cm\cubed} for the density and \SI{10000}{\K} for the temperature.

The solar wind magnetic field strength is in the range \SIrange{0.4}{62}{nT}, the velocity in the range \SIrange{156}{1189}{\km\per\s}, the density in the range \SIrange{0}{117}{\per\cm\cubed}, and the temperature in the range \SIrange{3450}{6.63e6}{\K}, the mean values are at \SI{6.28}{\nT}, \SI{436}{\km\per\s}, \SI{6.8}{\per\cm\cubed} and \SI{1.05e5}{\K}. These ranges and mean values are as statistically expected from previous analyses of near \SI{1}{\au} solar wind data (e.g., Table~3.3 in \citet[p.~39]{Bothmer2007}).

Much higher or lower peak values at \SI{1}{\au} have been observed in extraordinary events, such as the 23~July 2012 ICME with a speed of over \SI{2000}{\km\per\s} and a peak field strength of about \SI{100}{\nT} that was observed by STEREO~A \citep{Russell2013} or the solar wind disappearance event observed in May 1999 with density values even down to \SI{0.2}{\per\cm\cubed} \citep{Lazarus2000}.\\

%subsection{Lognormal fitting}
%choice of distribution function
The frequency distributions of the solar wind parameters magnetic field strength, proton density and temperature can
well be approximated by lognormal distributions, whereas the proton velocity’s frequency has a differing shape, as shown in \citet{Veselovsky2010}. We investigate how well all four solar wind parameters’ frequency distributions can be represented by lognormal functions, which we use in the process of a least squares regression fitting. The lognormal function 

@@

depends on the location $\mu$ and the shape parameter $\sigma$. Changes in $\mu$ affect both the horizontal and vertical scaling of the function whereas $\sigma$ influences its shape. The distribution's median $x_\text{med}$ and mean $x_\text{avg}$ (average) positions are straightforward to interprete and are directly calculated from $\mu$ and $\sigma$:

@@

The values of $x_\text{med}$ and $x_\text{avg}$ obtained from fitting the individual solar wind frequency distributions are listed in Table~\ref{tab:lognormal_fit_parameters}.

@@

From visual inspection, the resulting curves describe the shape of the magnetic field strength, density and temperature distributions well as shown in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}. However, for the velocity the fit function appears not to be as good in describing the measured distribution’s more complex shape around its maximum and in the higher velocity range. This also can be inferred from the sum of absolute residuals (SAR) between data and fit listed in Table~\ref{tab:lognormal_fit_parameters}, being almost three times larger than those from the other parameters.
\begin{figure*}
	\includegraphics[width=18cm]{figures/histogram_fits_4_a_zoom_paper_pdfplot.pdf}
	\caption{Frequency distributions of the four solar wind parameters and their lognormal fits. The histograms have bins of 0.5~nT, 10~km/s, 1~cm$^{-3}$ and 10\,000~K and are based on the hourly OMNI data set. The fit's median and mean values are indicated as well. The insets only have zoomed-in frequency axes, their x-axes stay the same.}
	\label{fig:histogram_fits_4_a_zoom_paper_pdfplot}
\end{figure*}

%double lognormal fitting
%%%%%%%%%%%%%%%%%%%%%%%%%
%justification for compositional lognormal fit
In order to find a better fit result for the velocity distribution we assume that the velocity distribution can be made up of at least two overlapping branches \citep{McGregor2011b}. Therefore a compositional approach  is chosen by combining two lognormal functions (\ref{eq:single_lognormal_fit_function}), involving more fit variables:
\begin{align}
	W_\text{II}(x) &= c \cdot W_1(x) + (1 -c) \cdot W_2(x)\,.	\label{eq:double_lognormal_fit_function}
\end{align}
The balancing parameter $c$ ensures that the resulting function remains normalized as it represents a probability distribution.

The fitting of $W_\text{II}(x)$ to the velocity's frequency distribution yields the values of the now five fit parameters ($c$, $x_\text{med,1}$, $x_\text{avg,1}$, $x_\text{med,2}$ and $x_\text{avg,2}$) as listed in Table~\ref{tab:lognormal_fit_parameters} together with the median and mean values of the composed distribution, which can be derived via solving
\begin{align}
	\int W_\text{II}(x)\,\text{d}x = 0	&	&\text{and}	&	&\int x\,W_\text{II}(x)\,\text{d}x = 0	\,.
\end{align}
\hl{XXXXXXXXXXXXXXXXXXXXX}\\
As anticipated, this more complex fit function is more accurate in describing the velocity's frequency distribution (see Fig.~\ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}).
\begin{figure}
	\resizebox{\hsize}{!}{\includegraphics{figures/histogram_fits_V_a_zoom_dbl_paper_pdfplot.pdf}}
	\caption{The velocity's frequency distribution (same as in Fig.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot}) and its compositional lognormal fit. The fit's median and mean values and its two fit parts are indicated as well. The inset only has a zoomed-in frequency axis, its x-axis stays the same.}
	\label{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}
\end{figure}
For this reason we keep using the double lognormal ansatz for the velocity frequency fits in the following sections.\\
In this static model the slow and fast part contribute almost equally ($c \approx 0.5$), which of course is only valid for this kind of long-term average. At different times in a solar cycle their contributions vary strongly.

@@

\subsection{Power law lognormal fitting}
To retrieve the frequency distributions for every \SI{0.01}{\au} solar distance bin, we choose the same solar wind parameter binning as with the OMNI data (Sect.~\ref{sec:frequency_distribution}).
% Helios histogram bin sizes for mean of frequency distribution (at specific solar distance):\\
% $B$: bin size 0.5~nT, min 337, mean precision: 0.000545\\
% $v$: bin size 1~cm$^{-3}$, min 497, mean precision: 0.00449\\
% $n$: bin size 10~km/s, min 497, mean precision:  0.00449\\
% $T$: bin size 10\,000~K, min 497, mean precision: 4.49--44.49\\
